---
title: "Estimating the Unobserved: A Simulation Study on Censoring and Truncation in Race Models of Choice and Response Time"
shorttitle: "Race Model Censoring and Truncation"
author:
  - name: Jeroen E. Timmerman
    corresponding: true
    orcid: 0009-0003-8208-0509
    email: j.e.timmerman@uva.nl
    affiliations:
      - name: University of Amsterdam
        department: Department of Psychology
        address: Nieuwe Achtergracht 129-B
        city: Amsterdam
        region: North-Holland
        postal-code: 1018 WS
abstract: "This is my abstract."
keywords: [censoring, truncation, missing data, evidence accumulation, choice response data, diffusion decision model, linear ballistic accumulator, Bayesian hierarchical modeling. decision making]
author-note:
  disclosures:
    conflict of interest: The author has no conflict of interest to declare.
documentmode: stu
course: "MI2324RM: Research Master's Internship"
professor: Andrew Heathcote
duedate: August 31st, 2024
floatsintext: true
format:
  html:
    toc: true
    comments:
      hypothesis: true
  apaquarto-docx: default
  apaquarto-pdf: default
bibliography: references.bib
notebook-links: false
license: "CC BY-NC"
execute:
  cache: true
---

```{r packages to load}
#| include: false

library(EMC2)
library(rtdists)
library(stringr)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(BayesFactor)
```

```{r ggplot_colors}
ft_colors <- c("#C400C4", "#0EBF0E")
```


```{r true parameters}
#| include: false

parsLBA <- c(log(2),3,1,log(2),log(.75),log(.2))
names(parsLBA) <- c("B", "v", "v_lMd", "A", "sv_lMd", "t0")
parsLBA_label <- c("B", "v", "v[true]", "A", "s[v[true]]", "t[0]") # for ggplot
parsRDM <- c(log(3),1,.4,log(.75),log(.2))
names(parsRDM) <- c("B", "v", "v_lMd", "s_lMd", "t0")
parsRDM_label <- c("B", "v", "v[true]", "s[true]", "t0")
parsLNR <- c(log(.75),log(.65),log(.5),log(.8),log(.4))
names(parsLNR) <- c("m", "m_lMd", "s", "s_lMd", "t0")
parsLNR_label <- c("m", "m[true]", "s", "s[true]", "t0")

parlabeller <- c(parsLBA_label, parsRDM_label, parsLNR_label)
names(parlabeller) <- c(names(parsLBA), names(parsRDM), names(parsLNR))
parlabeller <- parlabeller[!duplicated(parlabeller)]

```

```{r upper_censoring files}
#| include: false

file_names <- list.files("01_simulation/upper_censoring/EMCs")
EMCs_upper <- data.frame()
for (i in file_names){
  load(paste0("01_simulation/upper_censoring/EMCs/", i))
  EMC <- data.frame(EMC = str_remove(i, ".RData"),
                    Censoring = str_detect(i, "sM"),
                    Model = str_extract(i, "LBA|LNR|RDM"),
                    Missing = str_extract(i, "70|90|975")
                    )
  EMCs_upper <- rbind(EMCs_upper, EMC)
}
EMCs_upper <- na.omit(EMCs_upper) # to remove empty sLBA
EMCs_upper$Censoring <- factor(EMCs_upper$Censoring, c(FALSE, TRUE), c("Truncated", "Censored"))
EMCs_upper$Model <- factor(EMCs_upper$Model)
EMCs_upper$Missing <- factor(EMCs_upper$Missing, c(975, 90, 70), c("2.5%", "10%", "30%"))
```

```{r parameters for upper censoring}
#| include: false

pars_upper <- list()
for (i in 1:nrow(EMCs_upper)) {
  pars_upper[[EMCs_upper$EMC[i]]] <-
    as.data.frame(summary(get(EMCs_upper$EMC[i], envir = globalenv()),
                          selection = "alpha")[[1]])
  EMCs_upper$RMSE[i] <- sqrt(mean((pars_upper[[EMCs_upper$EMC[i]]][,2] - get(paste0("pars", EMCs_upper$Model[i]), envir = globalenv()))^2))
}

```

```{r CI for upper_censoring}
#| include: false
parameters_CI_upper <- do.call(rbind, pars_upper)
for (i in 1:nrow(parameters_CI_upper)) {
  name <- row.names(parameters_CI_upper)[i]
  parameters_CI_upper$Censoring[i] <- str_detect(name, "sM")
  parameters_CI_upper$Model[i] <- str_extract(name, "LBA|LNR|RDM")
  parameters_CI_upper$Missing[i] <- str_extract(name, "70|90|975")
  parameters_CI_upper$Parameter[i] <- sub(".*\\.", "", name)
  parameters_CI_upper$True[i] <- 
    get(paste0("pars", parameters_CI_upper$Model[i]), envir = globalenv())[parameters_CI_upper$Parameter[i]]
}

parameters_CI_upper$Missing <- factor(parameters_CI_upper$Missing, c(975, 90, 70), c("2.5%", "10%", "30%"))
parameters_CI_upper$Censoring <- factor(parameters_CI_upper$Censoring, c(TRUE, FALSE), c("Censored", "Truncated"))

```

```{r Wasserstein_upper}
#| include: false

Wasserstein_upper <- data.frame()

for (i in 1:nrow(EMCs_upper)) {
  MCMC_sample <- get_pars(get(EMCs_upper$EMC[i], envir = globalenv()),
                          return_mcmc = F, merge_chains = T)
  true_pars <- get(paste0("pars", as.character(EMCs_upper$Model[i])))
  W <- c(model = sqrt(mean(sqrt(colSums((MCMC_sample[,1,] - true_pars)^2)^2)),
         sqrt(rowMeans((MCMC_sample[,1,] - true_pars)^2)))
  W_column <- data.frame(W = W, Variable = names(W))
  W_addition <- cbind(EMCs_upper[i,], W_column, row.names = NULL)
  Wasserstein_upper <- rbind(Wasserstein_upper, W_addition)
}

```

```{r censoring_both files}
#| include: false

file_names_both <- list.files("01_simulation/censoring_both/EMCs")
EMCs_both <- data.frame()
for (i in file_names_both){
  load(paste0("01_simulation/censoring_both/EMCs/", i))
  EMC <- data.frame(EMC = str_remove(i, ".RData"),
                    Censoring = str_detect(i, "sM"),
                    Model = str_extract(i, "LBA|LNR|RDM"),
                    Tail = str_extract(i, "lower|upper|both"),
                    Missing = str_extract(i, "2|10|30|50"),
                    Response = !str_detect(i, "unknown")
                    )
  assign(EMC$EMC, s)
  EMCs_both <- rbind(EMCs_both, EMC)
}

EMCs_both$Censoring <- factor(EMCs_both$Censoring, c(FALSE, TRUE),
                         c("Truncated", "Censored"))
EMCs_both$Model <- factor(EMCs_both$Model)
EMCs_both$Response <- factor(EMCs_both$Response, c(FALSE, TRUE),
                              c("unknown", "known"))

```

```{r parameters for censoring_both}
#| include: false

pars_EMCs_both <- data.frame()
for (i in 1:nrow(EMCs_both)) {
  summ <- do.call(rbind, summary(get(EMCs_both$EMC[i]), by_subject = T))
  pars_EMCs_both <-
    rbind(
      pars_EMCs_both,
      data.frame(
        EMC = EMCs_both$EMC[i],
        Model = EMCs_both$Model[i],
        Censoring = EMCs_both$Censoring[i],
        Tail = EMCs_both$Tail[i],
        Missing = EMCs_both$Missing[i],
        Response = EMCs_both$Response[i],
        id = rep(1:10, each = nrow(summ) / 10),
        Variable = row.names(summ),
        True = get(paste0("pars", EMCs_both$Model[i]))[row.names(summ)],
        Q2.5 = summ[, 1],
        Median = summ[, 2],
        Q97.5 = summ[, 3],
        Rhat = summ[, 4],
        ESS = summ[, 5]
      )
    )
}

pars_EMCs_both$Missing <- factor(pars_EMCs_both$Missing,
                                     c(2,10,30,50),
                                     c("2%","10%", "30%", "50%"))

```

```{r Wasserstein_both}
#| include: false

Wasserstein_both <- data.frame()

for (i in 1:nrow(EMCs_both)) {
  MCMC_sample <- get_pars(get(EMCs_both$EMC[i], envir = globalenv()),
                          return_mcmc = F, merge_chains = T)
  true_pars <- get(paste0("pars", as.character(EMCs_both$Model[i])))
  W <- rbind(model = sqrt(rowMeans(sqrt(colSums((MCMC_sample[,1:10,] - true_pars)^2)^2))),
           sqrt(apply((MCMC_sample[,1:10,] - true_pars)^2, c(1,2), mean)))
  W_column <- data.frame(W = c(W), Variable = rownames(W)[row(W)],
                         id = colnames(W)[col(W)])
  W_addition <- cbind(EMCs_both[i,], W_column,
                                     row.names = NULL)
  Wasserstein_both <- rbind(Wasserstein_both, W_addition)
}

Wasserstein_both$Missing <- factor(Wasserstein_both$Missing,
                                     c(2,10,30,50),
                                     c("2%","10%", "30%", "50%"))

```


```{r RMSEs for censoring_both}
#| include: false
# Calculate RMSE per group
EMCs_both_RMSEs <- pars_EMCs_both %>%
  group_by(id, Missing, Response, 
           Tail, Censoring, Model) %>%
  summarise(
    RMSE = sqrt(mean((Median - True)^2)),  # Calculate RMSE
    .groups = 'drop'  # Ungroup after summarizing
  )
```

```{r MAEs for censoring_both}
#| include: false
# Calculate MAE per group
EMCs_both_MAEs <- pars_EMCs_both %>%
  group_by(id, Missing, Response, 
           Tail, Censoring, Model) %>%
  summarise(
    MAE = mean(abs(Median - True)),  # Calculate MAE
    .groups = 'drop'  # Ungroup after summarizing
  )
```

```{r correlations for censoring_both}
#| include: false
# Calculate R per group
EMCs_both_Rs <- pars_EMCs_both %>%
  group_by(id, Missing, Response, 
           Tail, Censoring, Model) %>%
  summarise(
    R = cor(Median, True),  # Calculate correlation
    .groups = 'drop'  # Ungroup after summarizing
  )
```

# Introduction {.content-hidden unless-format="html"}

Many paradigms in experimental psychology investigate speeded decision-making, or are adapted to be speeded decision-making tasks. There are two main outcome variables to these tasks: what choice someone made (or whether this matches the corresponding stimulus), and how fast someone made their choice. Researchers are often interested in the conditions that affect these decisions and response times (RTs).

A problem arises when we want to make comparisons about the performance on a task. Someone might be quicker to respond -- indicating better performance, but at the same time they might be less accurate -- indicating worse performance. This is commonly referred to as the speed accuracy trade-off, which complicates inferences on task performance.

A wide range of evidence accumulation models (EAMs) aim to model the cognitive processes behind decision making as noisy accumulation of evidence until a decision threshold is reached. This means that how fast a participant is able to accumulate evidence towards a choice (the evidence accumulation rate or drift rate) is modeled separately from a participants’ tendency to value speed over accuracy or vice versa (the threshold). In addition to these decisional variables, EAMs often estimate the non-decision time (the time it takes to encode the stimulus and the time it takes to make the motor response for example), and they often account for the variability within and between trials, as well as response bias.

This paper will focus on three prominent EAMs that are supported by the EMC2 package [@EMC2]: the Linear Ballistic Accumulator [LBA; @LBA], the Racing Diffusion Model [RDM; @RDM], and the Log-Normal Race [LNR; @LNR]. These models are race models, which model separate accumulators--or racers--for each choice option, with the first racer to hit the threshold reflecting the choice made. The time that it takes for the winning racer to hit the threshold is the decision time, which makes up the RT together with the non-decision time. 

The LBA [@LBA] models RTs and responses as a race between deterministic linear accumulators, with the slope for a racer drawn from $\mathcal{N}(v, s_{v}^2)$, where $v$ is the mean evidence accumulation rate and $s_v^2$ the between-trial variance of the evidence accumulation rate. The intercept also has between-trial variation and is drawn from $\mathcal{U}(0,A)$, the response threshold $b = B + A$ (meaning that B is the distance from the starting point variability to the response threshold), and the non-decision time is $t_0$. The line that intersects the threshold at the lowest time determines the decision made, and the timepoint of the intersection added to $t_0$ is the RT [@LBA]. The RDM [@RDM] has a similar set of parameters, but instead of having between-trial variation in evidence accumulation rate, the RDM has continuous normally distributed variation within each trial. Lastly, the LNR is a race between random samples from log-normal distributions, with the lowest sample determining the choice and decision time, which is then added to non-decision time $t_0$ to constitute the RT. This means that the LNR has three main parameters: the scale $m$ of the lognormal, the shape $s$ of the lognormal, and the non-decision time $t_0$ [@LNR]. @fig-racemodels illustrates the dynamics of each model, with A set to 0 for the RDM to reflect the parameters used in this paper.

```{r Race Models}
#| label: fig-racemodels
#| fig-cap: Simple Race Model Illustrations
#| apa-note: The dynamics of the race models in this paper are illustrated here. The Linear Ballistic Accumulator (LBA) accumulates evidence in a linear deterministic fashion, with normally distributed variability $s_v^2$ in slope $v$ and uniformly distributed variability $A$ in starting point between trials. The Racing Diffusion Model (RDM) has continuous normally distributed variance $s_v^2$ within the trial instead (diffusion), and can include starting point variability between trials like the LBA, although this is not the case for simulations in this paper. The Log-Normal Race (LNR) samples accumulation times from log-normal distributions, with the shortest accumulation time (solid vertical lines) determining the choice and the decision time.

# LBA ----

LBA_accumulator_plot <- ggplot(data.frame(RT = 0:2)) +
  geom_segment(aes(x = 0.2, y = 1, xend = (4 - 1) / 5 , yend = 4,
                   color = "True"),
               linewidth = 0.8,
               arrow = arrow(length = unit(3, "mm"), type = "open")) +
  geom_segment(aes(x = 0.2, y = 1.3, xend = (4 - 1.3) / 2.5 , yend = 4,
                   color = "False"),
               linewidth = 0.8,
               arrow = arrow(length = unit(3, "mm"), type = "open")) +
  theme_classic() +
  theme(plot.margin = margin(t = 0),
        axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_line(arrow = arrow(length = unit(4, "mm"),
                                                 type = "open")),
        legend.position = "none") +
  xlab("Response Time") +
  geom_hline(yintercept = 4, linewidth = 2) +
  geom_vline(xintercept = 0.001,
             linetype = "dashed") +
  geom_vline(xintercept = 0.2,
             linetype = "dashed") +
  geom_segment(x = 0.01, xend = 0.19, y = 3.3, yend = 3.3, linewidth = 0.6) +
  geom_segment(x = 0.01, xend = 0.01, y = 3.2, yend = 3.4) +
  geom_segment(x = 0.19, xend = 0.19, y = 3.2, yend = 3.4) +
  annotate("text", x = 0.1, y  = 3.1, label = "t[0]", parse = TRUE) +
  annotate("rect", xmin = 0.1, xmax = 0.2, ymin = 0, ymax = 2, alpha = 0.2) +
  annotate("text", x = 0.15, y = 1, label = "A") +
  geom_segment(x = 1.8, xend = 1.8, y = 0, yend = 4, 
               arrow = arrow(length = unit(3, "mm"), type = "open")) +
  annotate("text", x = 1.87, y = 2, label = "b") +
  scale_x_continuous(limits=c(0, 2), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_manual(values = ft_colors)

LBA_density_plot <- ggplot(data.frame(RT = 0:2)) +
  stat_function(fun = dLBA,
                aes(color = "True"),
                args = list(response = 2, A=2, b= 4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75), LT = LT, UT = UT),
                linewidth = 0.8) +
  stat_function(fun = dLBA,
                aes(color = "False"),
                args = list(response = 1, A=2, b= 4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75), LT = LT, UT = UT),
                linewidth = 0.8) +
  theme_classic() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = margin(b = 0),
        axis.ticks.length = unit(0, "pt"),
        axis.line.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits=c(0, 2), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("LBA") +
  scale_color_manual(values = ft_colors)

LBA_plot <- LBA_density_plot / LBA_accumulator_plot

# RDM ----
parsRDM <- c(log(3),1,.4,log(.75),log(.2))
names(parsRDM) <- c("B", "v", "v_lMd", "s_lMd", "t0")

set.seed(2)
dt = 0.001
accumulator_true <- 0
drift_true <- 1.4
s_true <- 0.75
t0 <- 0.2

while (accumulator_true[length(accumulator_true)] < 3) {
  accumulator_true <- append(accumulator_true,
                             accumulator_true[length(accumulator_true)] + 
                               dt * drift_true + rnorm(1, 0, sqrt(s_true^2 * dt)))
}

accumulator_false <- 0
drift_false <- 1
s_false <- 1

while (accumulator_false[length(accumulator_false)] < 3) {
  accumulator_false <- append(accumulator_false,
                             accumulator_false[length(accumulator_false)] + 
                               dt * drift_false + rnorm(1, 0, sqrt(s_false^2 * dt)))
}

RT_true <- t0 + seq(0, (length(accumulator_true) - 1) * dt, by = dt)
RT_false <- t0 + seq(0, (length(accumulator_false) - 1) * dt, by = dt)
RDM_dat <- data.frame(X = c(accumulator_true, accumulator_false), RT = c(RT_true, RT_false),
           Accumulator = c(rep("True", length(accumulator_true)),
                           rep("False", length(accumulator_false))))

RDM_accumulator_plot <- ggplot(RDM_dat, aes(RT, X, color = Accumulator)) + 
  geom_line() +
  theme_classic() +
  theme(plot.margin = margin(t = 0, l = 8),
        axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none") +
  xlab("Response Time") +
  scale_x_continuous(limits=c(0, 4), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  geom_segment(x = 0, xend = 4, y = 0, yend = 0, color = "black",
               arrow = arrow(length = unit(4, "mm"), type = "open")) +
  geom_hline(yintercept = 3, linewidth = 2) +
  geom_vline(xintercept = 0.001,
             linetype = "dashed") +
  geom_vline(xintercept = 0.2,
             linetype = "dashed") +
  geom_segment(x = 0.01, xend = 0.19, y = 2.4, yend = 2.4, linewidth = 0.6, 
               color = "black") +
  geom_segment(x = 0.01, xend = 0.01, y = 2.3, yend = 2.5,
               color = "black") +
  geom_segment(x = 0.19, xend = 0.19, y = 2.3, yend = 2.5,
               color = "black") +
  annotate("text", x = 0.1, y  = 2.2, label = "t[0]", parse = TRUE) +
  geom_segment(x = 3, xend = 3, y = 0, yend = 3, 
               arrow = arrow(length = unit(3, "mm"), type = "open"),
               color = "black") +
  annotate("text", x = 3.13, y = 1.5, label = "b") +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = ft_colors)

dRDM <- function(t, v = 1.4,  B = 3, t0 = 0.2) {
  t <- t - t0
  out <- numeric(length(t))
  out[t <= 0] <- 0
  out[t > 0] <- B/sqrt(2*pi*t[t>0]^3) * 
    exp(-(1/2) * (v * t[t>0] - B)^2 / t[t>0])
  out
}

pRDM <- function(t, v = 1,  B = 3, t0 = 0.2) {
  t <- t - t0
  out <- numeric(length(t))
  out[t <= 0] <- 0
  out[t > 0] <- pnorm((v*t[t>0] - B)/sqrt(t[t>0])) + 
                exp(2*v*B)*pnorm((-v*t[t>0] - B) / sqrt(t[t>0]))
  out
}

gRDM <- function(t, v1 = 1.4, v2 = 1,  B = 3, t0 = 0.2) {
  dRDM(t, v1, B, t0) * (1 - pRDM(t, v2, B, t0))
}

RDM_density_plot <- ggplot(data.frame(RT = 0:6)) +
  stat_function(fun = gRDM,
                args = list(v1 = drift_true, v2 = drift_false,  B = 3, t0 = t0),
                aes(color = "True"),
                linewidth = .8) +
  stat_function(fun = gRDM,
                args = list(v1 = drift_false, v2 = drift_true,  B = 3, t0 = t0),
                aes(color = "False"),
                linewidth = .8) + 
  theme_classic() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = margin(b = 0),
        axis.ticks.length = unit(0, "pt"),
        axis.line.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  scale_x_continuous(limits=c(0, 5), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("RDM") +
  scale_color_manual(values = ft_colors)


RDM_plot <- RDM_density_plot / RDM_accumulator_plot

# LNR ----

parsLNR <- c(log(.75),log(.65),log(.5),log(.8),log(.4))
names(parsLNR) <- c("m", "m_lMd", "s", "s_lMd", "t0")

LNR_true <- function(x){
  x <- x - 0.4
  dlnorm(x, log(0.75) + log(0.65), 0.5 + 0.8) * 
    (1 - plnorm(x, log(0.75), 0.5))
}

LNR_false <- function(x){
  x <- x - 0.4
  dlnorm(x, log(0.75), 0.5) * (1 - plnorm(x, log(0.75) + log(0.65), 0.5 + 0.8))
}

set.seed(1312)
sample_true <- rlnorm(50,  log(0.75) + log(0.65), 0.5 + 0.8)
sample_false <- rlnorm(50,  log(0.75), 0.5)
LNR_sample <- data.frame(RT = c(sample_true, sample_false) + 0.4, 
                         Choice = rep(c("Correct", "Incorrect"), each = 50),
                         Won = c(sample_true > sample_false, sample_true < sample_false))

LNR_plot <- ggplot(LNR_sample, aes(RT, color = Choice, linetype = Won)) +
  stat_function(fun = LNR_true,
                aes(color = "Correct"),
                linewidth = 0.8) +
  stat_function(fun = LNR_false,
                aes(color = "Incorrect"),
                linewidth = 0.8) +
  geom_segment(aes(x = RT, xend = RT), y = -0.3, yend = 0) + 
  geom_segment(x = 0, xend = 2.5, y = 0, yend = 0, color = "black",
               arrow = arrow(length = unit(4, "mm"), type = "open")) +
  geom_segment(x = 0, xend = 0.4, y = -0.1, yend = -0.1, color = "black") +
  geom_segment(x = 0, xend = 0, y = -0.07, yend = -0.13, color = "black") +
  geom_segment(x = 0.4, xend = 0.4, y = -0.07, yend = -0.13, color = "black") +
  annotate("text", x = 0.2, y = -0.2, label = "t[0]", parse = T) +
  xlim(0, 2.5) +
  scale_y_continuous(limits = c(-0.3, 1.5), expand = c(0,0)) +
  ggtitle("LNR") + 
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_blank()) +
  labs(y = "Response Time") +
  scale_color_manual(values = rev(ft_colors)) +
  guides(linetype = "none")

(LBA_plot | RDM_plot) / free(LNR_plot)

```

Each of these race models have defined probability density functions $p(t \mid \theta)$ and cumulative density functions $P(t \mid \theta)$ for the finishing time of a single accumulator. To compute the likelihood of parameter vector $\theta$ for a trial, we can take the probability density at that trial's RT for the parameter vector $\theta_i$ corresponding to the choice $i$, and multiply it by the probability of the other accumulators $j$ not finishing before accumulator $i$:
$$
\mathcal{L}(\theta | t, i) = p(t \mid \theta_{i}) \prod_{j \neq i}{1 - P(t \mid \theta_{j})},
$$ {#eq-race-likelihood}
which is the "defective distribution" for $i$, meaning that it integrates to the probability of response $i$.

Complicating the modelling of speeded decision making, RTs and choices are often missing on certain trials due to experimental design. A researcher may want to limit slow RTs in their experiment design, for example to reduce slow type II thinking [@dualprocess], or to emphasise speed. Alternatively, researchers may want to remove outlying responses that cannot have come from the process of interest (e.g., a response 0.05 seconds after stimulus onset, which is too fast for a decision making process to occur).

There are two main ways to handle missing values: truncation, which discards missing values with no assumption of the underlying distribution, and censoring, which assumes the proportion of missing values to reflect the true distribution, and takes this into account in the model estimation.

Although truncation could potentially improve parameter estimates by eliminating irrelevant outliers, outlier removal often increases estimation bias by excluding extreme but valid RTs [@miller; @ulrichmiller; @outliersRatcliff; @MLcensoring]. When truncation is used for missing values due to response windows, this could even exclude less extreme values, leading to more estimation bias. @fig-LBAtrunc illustrates how upper truncation might distort parameter estimation in a simple two forced choice decision task with slow errors. Trials with slower accumulation rates--which tend to result in more incorrect trials--get discarded, while trials with higher accumulation rates are recorded.

```{r LBA plot}
#| label: fig-LBAtrunc
#| fig-cap: Missing Data in the LBA
#| apa-note: 

LT = 0
UT = 1
l = 0.8

ntrials <- 40

set.seed(4)
slopes1 <- rnorm(ntrials,3,1)
intercepts1 <- runif(ntrials,0,2)
boundaryx1 <- (4 - intercepts1) / slopes1

slopes2 <- rnorm(ntrials,4,1.75)
intercepts2 <- runif(ntrials,0,2)
boundaryx2 <- (4 - intercepts2) / slopes2

won <- boundaryx1 < boundaryx2
late1 <- boundaryx1 > UT - 0.2
late2 <- boundaryx2 > UT - 0.2

accumulators <- data.frame(slopes1, intercepts1, boundaryx1,
                           slopes2, intercepts2, boundaryx2,
                           won, late1, late2)

LBA_densities <- ggplot(data.frame(rt = seq(0, 2, length.out = 1000)), aes(rt)) +
  stat_function(fun = dLBA,
                aes(color = "Incorrect"),
                args = list(response = 1, A=2, b= 4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75), LT = LT, UT = UT),
                xlim = c(0, UT+0.0001),
                n = 1001,
                linewidth = l) +
  stat_function(fun = dLBA,
                args = list(response = 2, A=2, b= 4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75), LT = LT, UT = UT),
                xlim = c(0, UT+0.0001),
                n = 1001,
                aes(color = "Correct"),
                linewidth = l) +
  stat_function(fun = dLBA,
                aes(color = "Incorrect"),
                args = list(response = 1, A=2, b=4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75)),
                xlim = c(UT, 2),
                linetype = "dashed",
                linewidth = l) +
  stat_function(fun = dLBA,
                args = list(response = 2, A=2, b=4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75)),
                xlim = c(UT, 2),
                aes(color = "Correct"),
                linetype = "dashed",
                linewidth = l) +
  stat_function(fun = dLBA,
                args = list(response = 1, A=2, b=4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75)),
                xlim = c(UT, 3),
                geom = "area",
                aes(fill = "Incorrect"),
                alpha = 0.25) +
  stat_function(fun = dLBA,
                args = list(response = 2, A=2, b=4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75)),
                xlim = c(UT, 2),
                geom = "area",
                aes(fill = "Correct"),
                alpha = 0.25) +
  geom_vline(xintercept = UT) + 
  theme_classic() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = margin(b = 0),
        axis.ticks.length = unit(0, "pt"),
        axis.line.y = element_blank(),
        axis.title.y = element_blank()) +
  scale_x_continuous(limits=c(0, 2), expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  guides(fill = "none") +
  labs(color = "Choice") +
  scale_color_manual(values = rev(ft_colors))

LBA_racers <- ggplot(data.frame(rt = seq(0, 2, length.out = 1000)), aes(rt)) +
  geom_segment(data = accumulators[won,],
               aes(x = 0.2, y = intercepts1, xend = boundaryx1 + 0.2, yend = 4,
                   linetype = late1, colour = "Incorrect")) +
  geom_segment(data = accumulators[!won,],
               aes(x = 0.2, y = intercepts2, xend = boundaryx2 + 0.2, yend = 4,
                   linetype = late2, colour = "Correct")) +
  scale_x_continuous(limits=c(0, 2), expand = c(0, 0)) +
  scale_y_continuous(limits=c(0, 4), expand = c(0, 0)) +
  theme_classic() +
  geom_hline(yintercept = 4, linewidth = 2) +
  geom_vline(xintercept = UT) +
  geom_vline(xintercept = 0.001,
             linetype = "dashed") +
  geom_vline(xintercept = 0.2,
             linetype = "dashed") +
  theme(plot.margin = margin(t = 0),
        axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.y = element_blank(),
        axis.line.x = element_line(arrow = arrow(length = unit(4, "mm"),
                                                 type = "open")),
        legend.position = "none") +
  xlab("Response Time") +
  geom_segment(x = 0.01, xend = 0.19, y = 3.8, yend = 3.8, linewidth = 0.6) +
  geom_segment(x = 0.01, xend = 0.01, y = 3.7, yend = 3.9) +
  geom_segment(x = 0.19, xend = 0.19, y = 3.7, yend = 3.9) +
  annotate("text", x = 0.1, y  = 3.6, label = "t[0]", parse = TRUE) +
  annotate("rect", xmin = 0.1, xmax = 0.2, ymin = 0, ymax = 2, alpha = 0.2) +
  annotate("text", x = 0.15, y = 2.2, label = "A") +
  geom_segment(x = 1.8, xend = 1.8, y = 0, yend = 4, 
               arrow = arrow(length = unit(3, "mm"), type = "open")) +
  annotate("text", x = 1.83, y = 2.5, label = "b") +
  scale_color_manual(values = rev(ft_colors))

LBA_densities / LBA_racers

```

Despite the issues, many researchers still regularly implement truncation, perhaps because of common practice, or because computing the likelihood for a censored RT often requires numerical integration of the likelihood function over the censored range, which can be slow to compute. For race models, this means that we can integrate @eq-race-likelihood over the censored time range $R$:
$$
\mathcal{L}(\theta | t \in R, i) = \int_{R} p(t \mid \theta_{i}) \prod_{j \neq i}{1 - P(t \mid \theta_{j})} \, dt.
$$ {#eq-censoring}

@eq-censoring can also be extended to accomodate missing responses in addition to missing RTs by summing @eq-censoring for each accumulator, resulting in the total probability of any response in time range $R$. Similarly, if we censored both fast and slow responses with no distinction between the two, we can sum the integrals over the lower and the upper range. We can even combine censoring and truncation with 
$$
\mathcal{L}(\theta | t \in R, i) \frac{\mathcal{L}(\theta | t \in [0, \infty), i)}{\mathcal{L}(\theta | t \in S, i)},
$$ {#eq-censoringandtruncation}
where $S$ is the range of all untruncated values.

Although censoring has been shown to result in more accurate parameter recovery than truncation in common RT distributions [@MLcensoring], this has not yet been established for race models. As data are routinely censored or truncated, this study aims to compare parameter recovery for different levels of censoring and truncation for the LBA, LNR, and RDM. We will examine differences between the models in sensitivity to censoring and truncation, and to what extent different levels of censoring and/or truncation cause estimation bias and/or imprecision in designs with varying numbers of trials.

Parameter estimates are expected to deteriorate fast with increased truncation, while increased censoring is expected to result in smaller and less biased deterioration of parameter recovery. Parameter posterior distributions are expected to be overconfident for truncation, whereas posterior distributions are expected to proportionally decrease in confidence with increased censoring.

# Methods

We compared RT censoring and truncation on parameter recovery in two simulation studies. The first simulation study compared upper censoring and truncation with responses known in a large number of samples to assess asymptotic parameter identifiability. The second simulation study assessed parameter recovery on a smaller number of trials to evaluate the practical differences between censoring and truncation, comparing a larger number of missing data scenarios. 

To assess parameter identifiability, we first simulated data using known parameters, which we then fit the same model to. This allows us to compare the known, "true" parameter values with our fitted parameter values. The code for all analyses and data generation are available on https://github.com/timmerj1/censoring-truncation-study-EAMs.

For both simulation studies, a simple model with two stimuli and two racers was used to generate the data. Conventional constants for the LBA's ($s_{v_{false}} = 1$) and the RDM's drift rate variance ($s_{false} = 1$) were used to generate and fit the data. Parameter values were chosen to reflect common RT and choice distributions in simple two forced choice tasks (see @tbl-pars for an overview of parameters used). For the LBA, decision thresholds were defined with $B = b - A$. As is the default in EMC2, parameters on the positive real line were log transformed. Data was generated using the `make_data` function from EMC2 [@EMC2].

::: {#tbl-pars .content-hidden unless-format="html" layout-nrow="3"}
| $B$ | $v$ | $v_{true}$ | $A$ | $s_{v_{true}}$ | $t_0$ |
|-----|-----|------------|-----|----------------|-------|
| 2   | 3   | 1          | 2   | 0.75           | 0.2   |

: LBA Parameters {#tbl-LBA-pars}

| $m$  | $m_{true}$ | $s$ | $s_{true}$ | $t_0$ |
|------|------------|-----|------------|-------|
| 0.75 | 0.65       | 0.5 | 0.8        | 0.4   |

: Log-Normal Race {#tbl-LNR-pars}

| $B$ | $v$ | $v_{true}$ | $s_{true}$ | $t_0$ |
|-----|-----|------------|------------|-------|
| 3   | 1   | 4          | 0.75       | .2    |

: Racing Diffusion Model {#tbl-RDM-pars}

Simulation Parameters\
_
:::

::: {.content-hidden when-format="html" apa-note="Parameters used to simulate response time and choice data for study 1 and 2. Note that parameters on the real positive line are estimated on a log scale in EMC2. Subscript 'true' relates to a match between stimulus and choice (a correct choice is made), and these parameters are added to the corresponding parameter for the non-matching racer."}
| LBA               |           | LNR             |           | RDM            |          |
|-------------------|-----------|-----------------|-----------|----------------|----------|
| Parameter         | Value     | Parameter       | Value     | Parameter      | Value    |
|-------------------|-----------|-----------------|-----------|----------------|----------|
| $B$               | 2         | $m$             | 0.75      | $B$            | 3        |
| $v$               | 3         | $m_{true}$      | 0.65      | $v$            | 1        |
| $v_{true}$        | 1         | $s$             | 0.5       | $v_{true}$     | 4        |
| $A$               | 2         | $s_{true}$      | 0.8       | $s_{true}$     | 0.75     |
| $s_{v_{true}}$    | 0.75      | $t_0$           | 0.4       | $t_0$          | 0.2      |
| $B$               | 2         |                 |           |                |          |

Simulation Parameters
:::

For the first simulation, upper censoring and truncation were compared using a large number of trials (20,000 trials) to investigate the differences between censoring and truncation without fits being affected by random error. RTs were cut off at three different levels: at 2.5%, 10%, and 30% of upper values, with cutoff points estimated in a separate simulation of 20,000 trials. Responses were not missing for any of the censored values. For each combination of model and missing level, new data was simulated, but censoring and truncation were compared on the same datasets to ensure differences cannot be caused by random sampling error.

In the second simulation, a factorial design was used to vary whether responses were known or unknown, which tail missed response times (lower, upper, or both tails), the percentage of missing responses (2%, 10%, 30%, or 50%), and whether missing responses were censored or truncated. For each condition, ten samples with a small number of trials (400 trials) were simulated and fit. Missing level cutoffs were chosen using the quantiles on a non-missing simulation with 4000 trials.

Parameter posteriors were sampled using the particle Metropolis within Gibbs (PMwG) sampler in EMC2 [@EMC2] using its default (standard normal) priors for non-hierarchical estimation. Three PMwG chains were sampled for each parameter, with 50 particles per parameter [@PMwG]. 

To assess parameter recovery, the medians, the 2.5, and 97.5 percentiles of the PMwG samples were computed and plotted along with the true parameters used to simulate the data. To assess the fits numerically, root mean square errors (RMSEs) were computed for the sample medians for each participant fit $\boldsymbol{\hat{\theta}}$ compared to the true parameters $\boldsymbol{\theta}$: $$
RMSE = \sqrt{\frac{1}{n}\sum_{i=1}^{n}{(\boldsymbol{\theta} - \boldsymbol{\hat{\theta}})^2}}
$$ To make sure that the RMSEs were not overestimated due to outliers, mean absolute errors (MAEs) were also compared: $$
MAE = \frac{1}{n}\sum_{i=1}^{n}{|\boldsymbol{\theta} - \boldsymbol{\hat{\theta}}|}
$$ with $n$ denoting the number of parameters. Lastly, to be able to compare between models, Pearson's correlation coefficients were computed for each fit.

# Results

# Study 1: Upper Censoring

@fig-RMSE-upper shows the RMSE between the true parameter values and the medians of the posterior distribution for each fit. As expected, the RMSE for the fits increase with an increase of percentage missing for truncation. For censoring however, the RMSE stays close to 0 and does not clearly increase when data is censored rather than truncated. This indicates that censoring did improve the parameter recovery in an asymptotic fit as expected. The only case where censoring does not seem to outperform truncation is for the RDM, where 2.5% truncation did not perform worse than 2.5% censoring. With higher missing percentages, censoring still outperformed truncation for the RDM.

```{r RMSE_plot}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

EMCs_upper %>%
  ggplot(aes(Censoring, RMSE, fill = interaction(Missing, Censoring, sep = " "))) +
  facet_grid( ~ Model) +
  scale_fill_brewer(palette = "PuOr") +
  geom_bar(position = "dodge", stat = "identity") +
  labs(fill = "Missing") +
  theme_classic() +
  xlab("Censoring")
```

When looking at @fig-upper-censoring-recoveries, the generally higher RMSE for RDMs is explained by a general tendency for $B$ and $v$ to be overestimated, while $v_{win}$ and $t_0$ are underestimated. Overall, these parameters still seem to be recovered better with censoring than with truncation, except that $t_0$ and $s_{win}$ were recovered slightly better for truncation at a low percentage in this simulation. For the other models, censoring clearly performs better than truncation, with true parameters included by most credible intervals for censoring, while most truncation credible intervals exclude the true parameters. The main exception to this is the $s_{win}$ parameter for the LNR, where the credible interval for 30% censoring does not include the true parameter value, but all other credible intervals do.

```{r}
#| include: true
#| label: fig-upper-censoring-recoveries
#| fig-cap: "Parameter Recovery Plots for Upper Censoring / Truncation"
#| fig-alt: ""
#| apa-note: "Race model recoveries separated by model (LBA, LNR and RDM) and censoring vs truncation. Especially the LBA and LNR parameters are on the identity line for censoring, but off the line for truncation. "

LBA_upper_CI <- parameters_CI_upper %>%
  filter(Model == "LBA") %>%
  ggplot(aes(`50%`, interaction(Missing, Censoring, sep = " "), color = Missing)) +
  geom_point() +
  geom_errorbarh(aes(xmin = `2.5%`, xmax = `97.5%`)) +
  facet_grid( ~ Parameter, scale = "free_x", labeller = labeller(.default = label_parsed, Parameter = parlabeller)) +
  xlab("Parameter Estimates") + 
  ylab("") +
  geom_vline(aes(xintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("LBA") +
  scale_y_discrete(limits=rev)

LNR_upper_CI <- parameters_CI_upper %>%
  filter(Model == "LNR") %>%
  ggplot(aes(`50%`, interaction(Missing, Censoring, sep = " "), color = Missing)) +
  geom_point() +
  geom_errorbarh(aes(xmin = `2.5%`, xmax = `97.5%`)) +
  facet_grid( ~ Parameter, scale = "free_x", labeller = labeller(.default = label_parsed, Parameter = parlabeller)) +
  xlab("Parameter Estimates") + 
  ylab("") +
  geom_vline(aes(xintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("LNR") +
  scale_y_discrete(limits=rev)

RDM_upper_CI <- parameters_CI_upper %>%
  filter(Model == "RDM") %>%
  ggplot(aes(`50%`, interaction(Missing, Censoring, sep = " "), color = Missing)) +
  geom_point() +
  geom_errorbarh(aes(xmin = `2.5%`, xmax = `97.5%`)) +
  facet_grid( ~ Parameter, scale = "free_x", labeller = labeller(.default = label_parsed, Parameter = parlabeller)) +
  xlab("Parameter Estimates") + 
  ylab("") +
  geom_vline(aes(xintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("RDM") +
  scale_y_discrete(limits=rev)

LBA_upper_CI / LNR_upper_CI / RDM_upper_CI

```

Overall, these results indicate that upper censoring results in better parameter recovery than upper truncation in an asymptotic sample, i.e. with a large number of trials. This cannot yet be generalized to smaller sample sizes, as random error and parameter tradeoffs might affect parameter recovery more than censoring versus truncation does. Moreover, one might want to use lower censoring or truncation instead, or a combination of lower and upper censoring or truncation. Lastly, in this simulation the responses were known. Often censoring or truncation is implemented when there is a response time window, where neither the RTs or the choices are recorded. To account for these issues, the second simulation study takes these factors into account by using a smaller number of trials (200 instead of 1000 trials), in a $2 \times 2 \times 3 \times 3 \times 4$ design: censoring versus truncation, known versus unknown, lower versus upper versus both tails missing, LBA versus LNR versus RDM, and 2%, 10%, 30%, and 50% censoring or truncation.

# Study 2 Upper or Lower Censoring With and Without Decision

## LBA

Contrary to the first study, upper censoring with responses known did not improve the parameter recovery for the LBA compared to upper truncation. @fig-RMSE-both-LBA shows a similar performance for upper censoring and truncation, with the exception of censoring having clear misfits in the case of higher missing percentages when responses are known. Interestingly, upper truncation resulted in RMSEs that are on par with censoring in the lower tail or both tails.

Lower censoring outperformed lower truncation in terms of RMSE, both when responses were known or unknown. This is not very pronounced at 2% missing, but at higher missing percentages the differences are clear. Censoring at both tails also seems to improve estimates,

```{r}
#| include: true
#| label: fig-RMSE-both-LBA
#| fig-cap: "RMSE for the linear ballistic accumulator (LBA)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_RMSEs %>%
#   filter(Model == "LBA") %>%
#   summarise(mean = mean(RMSE), RMSE_CI = sd(RMSE) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - RMSE_CI, ymax = mean + RMSE_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_RMSEs %>%
  filter(Model == "LBA") %>%
  ggplot(aes(Missing, RMSE, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-MAE-both-LBA
#| fig-cap: "MAE for the linear ballistic accumulator (LBA)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_MAEs %>%
#   filter(Model == "LBA") %>%
#   summarise(mean = mean(MAE), 
#             MAE_CI = sd(MAE) / sqrt(n()),
#             .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - MAE_CI, ymax = mean + MAE_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_MAEs %>%
  filter(Model == "LBA") %>%
  ggplot(aes(Missing, MAE, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-R-both-LBA
#| fig-cap: "R for the linear ballistic accumulator (LBA)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_Rs %>%
#   filter(Model == "LBA") %>%
#   summarise(mean = mean(R),
#             R_CI = sd(R) / sqrt(n()),
#             .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - R_CI, ymax = mean + R_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_Rs %>%
  filter(Model == "LBA") %>%
  ggplot(aes(Missing, R, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-RMSE-both-LBA
#| fig-cap: "RMSE for the linear ballistic accumulator (LBA)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_Distances %>%
#   filter(Model == "LBA") %>%
#   summarise(mean = mean(Distance), 
#             Distance_CI = sd(Distance) / sqrt(n()), 
#             .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - Distance_CI, ymax = mean + Distance_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_Distances %>%
  filter(Model == "LBA") %>%
  ggplot(aes(Missing, Distance, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r Wasserstein_plot LBA}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

W_plots <- list()
for (i in names(parsLBA)) {
  W_plots[[i]] <- Wasserstein_both %>%
    filter(Model == "LBA") %>%
    filter(Variable == i) %>%
    ggplot(aes(Missing, Wasserstein_Distance, color = Censoring)) +
    scale_fill_brewer(palette = "PuOr") +
    facet_grid(Response ~ Tail,
             labeller = label_both) +
    geom_point(position = position_jitter(width = 0.3, height = 0)) +
    labs(fill = "Missing") +
    theme_classic() +
    xlab("Censoring") + 
    ylab("W") +
    ggtitle(parse(text = parlabeller[i]))
}

wrap_plots(W_plots, nrow = 3) + plot_layout(guides = "collect")

```


```{r}
#| include: true
#| label: fig-recoveries
#| fig-cap: "recoveries"
#| apa-note: "A note about the recoveries"

pars_EMCs_both$id <- as.factor(pars_EMCs_both$id)
pars_EMCs_both %>%
  filter(Model == "LBA" & Tail == "upper" & Response == "known") %>%
  ggplot(aes(Median, id, color = Missing)) +
  scale_color_viridis_d(direction = -1, option = "D") +
  geom_point() +
  facet_grid(Censoring + Missing ~ Variable, scales = "free_x", 
             labeller = labeller(.rows = label_value, Variable = parlabeller, 
                                 .default = label_parsed)) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5)) +
  geom_vline(aes(xintercept = True), linetype = "dashed") +
  theme_classic() + 
  xlab("Parameter Estimates (95% Credible Interval)") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        strip.text.x = element_text(size = 12))

```

```{r}
#| include: true
#| label: fig-recoveries
#| fig-cap: "recoveries"
#| apa-note: "A note about the recoveries"

pars_EMCs_both %>%
  filter(Model == "LBA" & Tail == "upper" & Response == "known") %>%
  ggplot(aes(Median, id, color = Missing)) +
  scale_color_viridis_d(direction = -1, option = "D") +
  geom_point() +
  facet_grid(Censoring + Missing ~ Variable, scales = "free_x", 
             labeller = labeller(.rows = label_value, Variable = parlabeller, 
                                 .default = label_parsed)) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5)) +
  geom_vline(aes(xintercept = True), linetype = "dashed") +
  theme_classic() + 
  xlab("Parameter Estimates (95% Credible Interval)") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        strip.text.x = element_text(size = 12))

```

```{r}

upper50known_burn10 <- get_pars(sMLBAupper50known, selection = "alpha",
                                stage = "burn", subject = 10)
upper50known_adapt10 <- get_pars(sMLBAupper50known, selection = "alpha",
                                stage = "adapt", subject = 10)
upper50known_sample10 <- get_pars(sMLBAupper50known, selection = "alpha",
                                 stage = "sample", subject = 10)
df_upper50known_burn10 <-  data.frame()
df_upper50known_adapt10 <-  data.frame()
df_upper50known_sample10 <- data.frame()

for (par in names(parsLBA)) {
  for (chain in 1:3) {
    df_upper50known_burn10 <- rbind(df_upper50known_burn10, 
                                     data.frame(Parameter = par, Chain = chain,
                                                Value = c(upper50known_burn10[[par]][[chain]]),
                                                Iteration = seq_along(
                                                  c(upper50known_burn10[[par]][[chain]]
                                                )),
                                                True = parsLBA[[par]]))
    df_upper50known_adapt10 <- rbind(df_upper50known_adapt10, 
                                     data.frame(Parameter = par, Chain = chain,
                                                Value = c(upper50known_adapt10[[par]][[chain]]),
                                                Iteration = seq_along(
                                                  c(upper50known_adapt10[[par]][[chain]]
                                                )),
                                                True = parsLBA[[par]]))
    df_upper50known_sample10 <- rbind(df_upper50known_sample10, 
                                     data.frame(Parameter = par, Chain = chain,
                                                Value = c(upper50known_sample10[[par]][[chain]]),
                                                Iteration = seq_along(
                                                  c(upper50known_sample10[[par]][[chain]])),
                                                True = parsLBA[[par]]))
  }
}

df_upper50known_burn10$Chain <- factor(df_upper50known_burn10$Chain)
df_upper50known_adapt10$Chain <- factor(df_upper50known_adapt10$Chain)
df_upper50known_sample10$Chain <- factor(df_upper50known_sample10$Chain)

df_upper50known_burn10 %>%
  ggplot(aes(x= Iteration, y = Value, color = Chain)) +
  facet_grid(Parameter ~ ., scales = "free_y", 
             labeller = labeller(Parameter = parlabeller, 
                                 .default = label_parsed)) +
  geom_line() +
  geom_hline(aes(yintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("Burn-In Stage")

df_upper50known_adapt10 %>%
  ggplot(aes(x= Iteration, y = Value, color = Chain)) +
  facet_grid(Parameter ~ ., scales = "free_y", 
             labeller = labeller(Parameter = parlabeller, 
                                 .default = label_parsed)) +
  geom_line() +
  geom_hline(aes(yintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("Adaptation Stage")

df_upper50known_sample10 %>%
  ggplot(aes(x= Iteration, y = Value, color = Chain)) +
  facet_grid(Parameter ~ ., scales = "free_y", 
             labeller = labeller(Parameter = parlabeller, 
                                 .default = label_parsed)) +
  geom_line() +
  geom_hline(aes(yintercept = True), linetype = "dashed") +
  theme_classic()+
  ggtitle("Sampling Stage")

```

```{r}

upper50known_burn7 <- get_pars(sMLBAupper50known, selection = "alpha",
                                stage = "burn", subject = 7)
upper50known_adapt7 <- get_pars(sMLBAupper50known, selection = "alpha",
                                stage = "adapt", subject = 7)
upper50known_sample7 <- get_pars(sMLBAupper50known, selection = "alpha",
                                 stage = "sample", subject = 7)
df_upper50known_burn7 <-  data.frame()
df_upper50known_adapt7 <-  data.frame()
df_upper50known_sample7 <- data.frame()

for (par in names(parsLBA)) {
  for (chain in 1:3) {
    df_upper50known_burn7 <- rbind(df_upper50known_burn7, 
                                     data.frame(Parameter = par, Chain = chain,
                                                Value = c(upper50known_burn7[[par]][[chain]]),
                                                Iteration = seq_along(
                                                  c(upper50known_burn7[[par]][[chain]]
                                                )),
                                                True = parsLBA[[par]]))
    df_upper50known_adapt7 <- rbind(df_upper50known_adapt7, 
                                     data.frame(Parameter = par, Chain = chain,
                                                Value = c(upper50known_adapt7[[par]][[chain]]),
                                                Iteration = seq_along(
                                                  c(upper50known_adapt7[[par]][[chain]]
                                                )),
                                                True = parsLBA[[par]]))
    df_upper50known_sample7 <- rbind(df_upper50known_sample7, 
                                     data.frame(Parameter = par, Chain = chain,
                                                Value = c(upper50known_sample7[[par]][[chain]]),
                                                Iteration = seq_along(
                                                  c(upper50known_sample7[[par]][[chain]])),
                                                True = parsLBA[[par]]))
  }
}

df_upper50known_burn7$Chain <- factor(df_upper50known_burn7$Chain)
df_upper50known_adapt7$Chain <- factor(df_upper50known_adapt7$Chain)
df_upper50known_sample7$Chain <- factor(df_upper50known_sample7$Chain)

df_upper50known_burn7 %>%
  ggplot(aes(x= Iteration, y = Value, color = Chain)) +
  facet_grid(Parameter ~ ., scales = "free_y", 
             labeller = labeller(Parameter = parlabeller, 
                                 .default = label_parsed)) +
  geom_line() +
  geom_hline(aes(yintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("Burn-In Stage")

df_upper50known_adapt7 %>%
  ggplot(aes(x= Iteration, y = Value, color = Chain)) +
  facet_grid(Parameter ~ ., scales = "free_y", 
             labeller = labeller(Parameter = parlabeller, 
                                 .default = label_parsed)) +
  geom_line() +
  geom_hline(aes(yintercept = True), linetype = "dashed") +
  theme_classic() +
  ggtitle("Adaptation Stage")

df_upper50known_sample7 %>%
  ggplot(aes(x= Iteration, y = Value, color = Chain)) +
  facet_grid(Parameter ~ ., scales = "free_y", 
             labeller = labeller(Parameter = parlabeller, 
                                 .default = label_parsed)) +
  geom_line() +
  geom_hline(aes(yintercept = True), linetype = "dashed") +
  theme_classic()+
  ggtitle("Sampling Stage")

```

```{r}

dadm <- sMLBAupper50known[[1]]$data$`10`

# Copied from EMC2:::calc_ll_manager() () ----

c_name <- attr(dadm,"model")()$c_name
p_types <- names(attr(dadm, "model")()$p_types)
designs <- list()
for (p in c(p_types, attr(attr(dadm, "adaptive"), "aptypes"))) {
  designs[[p]] <-
    attr(dadm, "designs")[[p]][attr(attr(dadm, "designs")[[p]], "expand"), ,
                               drop = FALSE]
}
constants <- attr(dadm, "constants")
if (is.null(constants))
  constants <- NA
n_trials = nrow(dadm)
if (c_name == "DDM") {
  levels(dadm$R) <- c(0, 1)
  pars <- get_pars(proposals[1, ], dadm)
  pars <- cbind(pars, dadm$R)
  parameter_char <- apply(pars, 1, paste0, collapse = "\t")
  parameter_factor <-
    factor(parameter_char, levels = unique(parameter_char))
  parameter_indices <-
    split(seq_len(nrow(pars)), f = parameter_factor)
  names(parameter_indices) <- 1:length(parameter_indices)
} else{
  parameter_indices <- list()
}

# calc_ll ----



betterstuff <- EMC2:::calc_ll(t(as.matrix(parsLBA)), dadm, constants = constants, designs = designs, type = c_name, p_types = p_types, min_ll = log(1e-10), group_idx = parameter_indices)

pars_false <- matrix(pars_EMCs_both$Median[pars_EMCs_both$EMC == "sMLBAupper50known" & pars_EMCs_both$id == 10], nrow = 1)
colnames(pars_false)<- names(parsLBA)

badstuff <- EMC2:::calc_ll(pars_false, dadm, constants = constants, designs = designs, type = c_name, p_types = p_types, min_ll = log(1e-10), group_idx = parameter_indices)

betterstuff
badstuff

```

```{r}

ggplot(data.frame(rt = seq(0, 2, length.out = 1000)), aes(rt)) +
  stat_function(fun = dLBA,
                aes(color = "Incorrect", linetype = "True"),
                args = list(response = 1, A=2, b= 4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75))) +
  stat_function(fun = dLBA,
                aes(color = "Correct", linetype = "True"),
                args = list(response = 2, A=2, b= 4, t0 = .2, mean_v=c(3,4), sd_v=c(1,1.75))) +
  stat_function(fun = dLBA,
                aes(color = "Incorrect", linetype = "Off"),
                args = list(response = 1, A=0.05, b= 0.1, t0 = .6, mean_v=c(-4,0), sd_v=c(1,2.2))) +
  stat_function(fun = dLBA,
                aes(color = "Correct", linetype = "Off"),
                args = list(response = 2, A=0.05, b= 0.1, t0 = .6, mean_v=c(-4,0), sd_v=c(1,2.2))) +
  scale_linetype_manual(breaks = c("True", "Off"), values = c(1, 2))

```

## LNR

```{r}
#| include: true
#| label: fig-RMSE-both-LNR
#| fig-cap: "RMSE for the linear ballistic accumulator (LNR)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_RMSEs %>%
#   filter(Model == "LNR") %>%
#   summarise(mean = mean(RMSE), RMSE_CI = sd(RMSE) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - RMSE_CI, ymax = mean + RMSE_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_RMSEs %>%
  filter(Model == "LNR") %>%
  ggplot(aes(Missing, RMSE, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-MAE-both-LNR
#| fig-cap: "MAE for the linear ballistic accumulator (LNR)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_MAEs %>%
#   filter(Model == "LNR") %>%
#   summarise(mean = mean(MAE), MAE_CI = sd(MAE) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - MAE_CI, ymax = mean + MAE_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_MAEs %>%
  filter(Model == "LNR") %>%
  ggplot(aes(Missing, MAE, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-R-both-LNR
#| fig-cap: "R for the linear ballistic accumulator (LNR)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_Rs %>%
#   filter(Model == "LNR") %>%
#   summarise(mean = mean(R), R_CI = sd(R) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - R_CI, ymax = mean + R_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_Rs %>%
  filter(Model == "LNR") %>%
  ggplot(aes(Missing, R, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r Wasserstein_plot LNR}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

W_plots <- list()
for (i in names(parsLNR)) {
  W_plots[[i]] <- Wasserstein_both %>%
    filter(Model == "LNR") %>%
    filter(Variable == i) %>%
    ggplot(aes(Missing, Wasserstein_Distance, color = Censoring)) +
    scale_fill_brewer(palette = "PuOr") +
    facet_grid(Response ~ Tail,
             labeller = label_both) +
    geom_point(position = position_jitter(width = 0.3, height = 0)) +
    labs(fill = "Missing") +
    theme_classic() +
    xlab("Censoring") + 
    ylab("W") +
    ggtitle(parse(text = parlabeller[i]))
}

wrap_plots(W_plots, nrow = 3) + plot_layout(guides = "collect")

```

## RDM

```{r}
#| include: true
#| label: fig-RMSE-both-RDM
#| fig-cap: "RMSE for the linear ballistic accumulator (RDM)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_RMSEs %>%
#   filter(Model == "RDM") %>%
#   summarise(mean = mean(RMSE), RMSE_CI = sd(RMSE) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - RMSE_CI, ymax = mean + RMSE_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_RMSEs %>%
  filter(Model == "RDM") %>%
  ggplot(aes(Missing, RMSE, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-MAE-both-RDM
#| fig-cap: "MAE for the linear ballistic accumulator (RDM)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_MAEs %>%
#   filter(Model == "RDM") %>%
#   summarise(mean = mean(MAE), MAE_CI = sd(MAE) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - MAE_CI, ymax = mean + MAE_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_MAEs %>%
  filter(Model == "RDM") %>%
  ggplot(aes(Missing, MAE, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| include: true
#| label: fig-R-both-RDM
#| fig-cap: "R for the linear ballistic accumulator (RDM)"
#| fig-alt: ""
#| apa-note: "Something something"

# w = 0.3
# 
# EMCs_both_Rs %>%
#   filter(Model == "RDM") %>%
#   summarise(mean = mean(R), R_CI = sd(R) / sqrt(n()), .by = c(Response, Tail, Missing, Censoring)) %>%
#   ggplot(aes(Missing, mean, color = Censoring, group = Censoring)) +
#   facet_grid(Response ~ Tail,
#              labeller = label_both) +
#   geom_point(position = position_dodge(width = w)) +
#   geom_path(position = position_dodge(width = w)) +
#   geom_errorbar(aes(ymin = mean - R_CI, ymax = mean + R_CI),
#                 position = position_dodge(width = w),
#                 width = 0.5) +
#   labs(fill = "Missing") +
#   theme_classic() +
#   scale_y_continuous(n.breaks = 3)

EMCs_both_Rs %>%
  filter(Model == "RDM") %>%
  ggplot(aes(Missing, R, color = Censoring, group = Censoring)) +
  facet_grid(Response ~ Tail,
             labeller = label_both) +
  geom_point(position = position_jitter(width = 0.3, height = 0)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r Wasserstein_plot RDM}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

W_plots <- list()
for (i in names(parsRDM)) {
  W_plots[[i]] <- Wasserstein_both %>%
    filter(Model == "RDM") %>%
    filter(Variable == i) %>%
    ggplot(aes(Missing, Wasserstein_Distance, color = Censoring)) +
    scale_fill_brewer(palette = "PuOr") +
    facet_grid(Response ~ Tail,
             labeller = label_both) +
    geom_point(position = position_jitter(width = 0.3, height = 0)) +
    labs(fill = "Missing") +
    theme_classic() +
    xlab("Censoring") + 
    ylab("W") +
    ggtitle(parse(text = parlabeller[i]))
}

wrap_plots(W_plots, nrow = 3) + plot_layout(guides = "collect")

```


# Discussion

- no hierarchical structure --> more effect of random error
- no contaminants --> favors censoring
- no combinations of censoring and truncation
- no missing direction on both tails

# Appendix {.appendix}

```{r Wasserstein_plot LBA}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

W_plots <- list()
for (i in names(parsLBA)) {
  W_plots[[i]] <- Wasserstein_upper %>%
    filter(Model == "LBA") %>%
    filter(Variable == i) %>%
    ggplot(aes(Censoring, Wasserstein_Distance, 
               fill = interaction(Missing, Censoring, sep = " "))) +
    scale_fill_brewer(palette = "PuOr") +
    geom_bar(position = "dodge", stat = "identity") +
    labs(fill = "Missing") +
    theme_classic() +
    xlab("Censoring") + 
    ylab("W") +
    ggtitle(parse(text = parlabeller[i]))
}

wrap_plots(W_plots) + plot_layout(guides = "collect")

```

```{r Wasserstein_plot LNR}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

W_plots <- list()
for (i in names(parsLNR)) {
  W_plots[[i]] <- Wasserstein_upper %>%
    filter(Model == "LNR") %>%
    filter(Variable == i) %>%
    ggplot(aes(Censoring, Wasserstein_Distance, 
               fill = interaction(Missing, Censoring, sep = " "))) +
    scale_fill_brewer(palette = "PuOr") +
    geom_bar(position = "dodge", stat = "identity") +
    labs(fill = "Missing") +
    theme_classic() +
    xlab("Censoring") + 
    ylab("W") +
    ggtitle(parse(text = parlabeller[i]))
}

wrap_plots(W_plots) + plot_layout(guides = "collect")

```

```{r Wasserstein_plot RDM}
#| include: true
#| label: fig-RMSE-upper
#| fig-cap: "Root Mean Squared Errors (RMSE) for Upper Censoring in Race Models"
#| apa-note: "For each model and each level of data missing, the RMSE was computed."

W_plots <- list()
for (i in names(parsRDM)) {
  W_plots[[i]] <- Wasserstein_upper %>%
    filter(Model == "RDM") %>%
    filter(Variable == i) %>%
    ggplot(aes(Censoring, Wasserstein_Distance, 
               fill = interaction(Missing, Censoring, sep = " "))) +
    scale_fill_brewer(palette = "PuOr") +
    geom_bar(position = "dodge", stat = "identity") +
    labs(fill = "Missing") +
    theme_classic() +
    xlab("Censoring") + 
    ylab("W") +
    ggtitle(parse(text = parlabeller[i]))
}

wrap_plots(W_plots) + plot_layout(guides = "collect")

```

```{r}
#| include: true
#| label: fig-recoveries
#| fig-cap: "recoveries"
#| apa-note: "A note about the recoveries"

pars_EMCs_both$id <- as.factor(pars_EMCs_both$id)
pars_EMCs_both %>%
  filter(Model == "LBA" & Tail == "upper" & Response == "known") %>%
  ggplot(aes(Median, id, color = Missing)) +
  scale_color_viridis_d(direction = -1, option = "D") +
  geom_point() +
  facet_grid(Censoring + Missing ~ Variable, scales = "free_x") +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5)) +
  geom_vline(aes(xintercept = True), linetype = "dashed") +
  theme_classic() + 
  xlab("Parameter Estimates (95% Credible Interval)") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

```

```{r}

pars_EMCs_both[pars_EMCs_both$Model == "LBA" & pars_EMCs_both$tail == "upper" & pars_EMCs_both$percentage == "50%" & pars_EMCs_both$censoring == "Censored",]

```




```{r}
#| label: fig-RMSE-both
#| fig-cap: "Mean of Root Mean Squared Errors in each condition"
#| apa-note: "For each model, the RMSE was computed over all subject fits."

EMCs_both_RMSEs %>%
  ggplot(aes(censoring, RMSE, fill = interaction(percentage, censoring, sep = " "))) +
  facet_grid(tail + response_known ~ Model) +
  scale_fill_brewer(palette = "PuOr") +
  stat_summary(position = "dodge", geom = "bar", fun = "mean") +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)
  
```

```{r}
#| label: fig-MAE-both
#| fig-cap: "Mean of Root Mean Squared Errors in each condition"
#| apa-note: "For each model, the MAE was computed over all subject fits."

EMCs_both_MAEs %>%
  ggplot(aes(censoring, MAE, fill = interaction(percentage, censoring, sep = " "))) +
  facet_grid(tail + response_known ~ Model) +
  scale_fill_brewer(palette = "PuOr") +
  stat_summary(position = "dodge", geom = "bar", fun = "mean") +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| label: fig-r-both
#| fig-cap: "Mean of Root Mean Squared Errors in each condition"
#| apa-note: "For each model, the R was computed over all subject fits."

EMCs_both_Rs %>%
  ggplot(aes(censoring, R, fill = interaction(percentage, censoring, sep = " "))) +
  facet_grid(tail + response_known ~ Model) +
  scale_fill_brewer(palette = "PuOr") +
  stat_summary(position = "dodge", geom = "bar", fun = "mean") +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3)

```

```{r}
#| label: fig-RMSE-both
#| fig-cap: "Mean of Root Mean Squared Errors in each condition"
#| apa-note: "For each model, the R was computed over all subject fits."

RMSE_LBA <- EMCs_both_RMSEs %>%
  filter(Model == "LBA") %>%
  summarise(mean = mean(RMSE), RMSE_CI = qt(.975, 10 * 8 - 8) * sd(RMSE) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - RMSE_CI, ymax = mean + RMSE_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("LBA")

RMSE_LNR <- EMCs_both_RMSEs %>%
  filter(Model == "LNR") %>%
  summarise(mean = mean(RMSE), RMSE_CI = qt(.975, 10 * 8 - 8) * sd(RMSE) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - RMSE_CI, ymax = mean + RMSE_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("LNR")

RMSE_RDM <- EMCs_both_RMSEs %>%
  filter(Model == "RDM") %>%
  summarise(mean = mean(RMSE), RMSE_CI = qt(.975, 10 * 8 - 8) * sd(RMSE) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - RMSE_CI, ymax = mean + RMSE_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("RDM")

RMSE_LBA / RMSE_LNR / RMSE_RDM

```

```{r}
#| label: fig-MAE-both
#| fig-cap: "Mean of Root Mean Squared Errors in each condition"
#| apa-note: "For each model, the R was computed over all subject fits."

MAE_LBA <- EMCs_both_MAEs %>%
  filter(Model == "LBA", percentage != "50%") %>%
  summarise(mean = mean(MAE), MAE_CI = qt(.975, 10 * 8 - 8) * sd(MAE) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - MAE_CI, ymax = mean + MAE_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("LBA")

MAE_LNR <- EMCs_both_MAEs %>%
  filter(Model == "LNR", percentage != "50%") %>%
  summarise(mean = mean(MAE), MAE_CI = qt(.975, 10 * 8 - 8) * sd(MAE) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - MAE_CI, ymax = mean + MAE_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("LNR")

MAE_RDM <- EMCs_both_MAEs %>%
  filter(Model == "RDM", percentage != "50%") %>%
  summarise(mean = mean(MAE), MAE_CI = qt(.975, 10 * 8 - 8) * sd(MAE) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - MAE_CI, ymax = mean + MAE_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("RDM")

MAE_LBA / MAE_LNR / MAE_RDM

```

```{r}
#| label: fig-r-both
#| fig-cap: "Mean of Root Mean Squared Errors in each condition"
#| apa-note: "For each model, the R was computed over all subject fits."

R_LBA <- EMCs_both_Rs %>%
  filter(Model == "LBA") %>%
  summarise(mean = mean(R), R_CI = qt(.975, 10 * 8 - 8) * sd(R) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - R_CI, ymax = mean + R_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("LBA")

R_RDM <- EMCs_both_Rs %>%
  filter(Model == "RDM") %>%
  summarise(mean = mean(R), R_CI = qt(.975, 10 * 8 - 8) * sd(R) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - R_CI, ymax = mean + R_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("RDM")

R_LNR <- EMCs_both_Rs %>%
  filter(Model == "LNR") %>%
  summarise(mean = mean(R), R_CI = qt(.975, 10 * 8 - 8) * sd(R) / sqrt(n()), .by = c(response_known, tail, percentage, censoring)) %>%
  ggplot(aes(percentage, mean, color = censoring)) +
  facet_grid(response_known ~ tail) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - R_CI, ymax = mean + R_CI)) +
  labs(fill = "Missing") +
  theme_classic() +
  scale_y_continuous(n.breaks = 3) +
  ggtitle("LNR")

R_LBA/ R_RDM / R_LNR

```
